---
title: "A quick intro to R"
output:
  pdf_document:
    toc: true
    number_sections: true
---

\newpage

# Load packages


```{r, warning=FALSE,  message = FALSE}
# install.packages("MASS") first if not already installed.
library(MASS) # Dataset
library(tidyverse) # Datamanipulation & plots
library(broom) # Functions to extract model statistics and parameters
library(stargazer) # Tables for statistical models
library(naniar) # Visualizing missing data
```

# Working in R/RStudio

# View data


Bemærk `tidyverse` (`dplyr`) "overskriver" en række funktioner fra pakkerne `stats` og `MASS`

This data frame contains the responses of 237 Statistics I students at the University of Adelaide to a number of questions.


```{r}
head(survey)
```
\pagebreak

# Datamanipulation

## Filter (Row-operations)


```{r}
survey %>% 
  filter(Smoke == "Never") %>% # R er case-sensitive
  head()
```
```{r}
survey %>% 
  filter(Pulse > 70) %>% 
  head()
```
Kombiner

```{r}
survey %>% 
  filter(Pulse > 70 & Smoke == "Never") %>% 
  head()
```

# Load and save data

TODO write_csv, read_csv

## Select (Column-operations)

```{r}
survey %>% 
  select(Fold:Clap) %>% 
  head()
```

```{r}
survey %>% 
  select(ends_with("Hnd")) %>% 
  head()
```
\pagebreak

# Visuzalize missing data


```{r}
vis_miss(survey)
```
\pagebreak

# Statistical modeling


## Remove missing

```{r}
estimation_data <- 
  survey %>% 
  select(-Pulse, -M.I, - Height) %>% # Remove columns
  filter(!if_any(everything(),
                    ~ is.na(.)
                 )) # Remove obs with any missing
estimation_data %>% head()
```

```{r}
estimation_data %>% 
  count(W.Hnd) %>% 
  mutate(share = n / sum(n))
```

## Visualize data

```{r}
estimation_data %>% 
  mutate(W.Hnd_numeric = W.Hnd %>% as.numeric() - 1 ) %>% # Make variable 0-based
  ggplot(aes(x = Age, y = W.Hnd_numeric)) +
  geom_point() +
  geom_smooth(method = "glm",
               method.args = list(family = "binomial"),
              se = FALSE)
```

## Run regression (logit)

```{r}
model1 <-
  glm(formula = W.Hnd ~ Sex + Fold + Clap + Exer + Smoke + Age,
    family = "binomial", 
    data = estimation_data
    )
model1 # default output
```

```{r}
model2 <-
  glm(formula = W.Hnd ~ Sex + Clap + Exer + Smoke + Age,
    family = "binomial", 
    data = estimation_data
    )
```

## Single row model summary

```{r}
glance(model1)
```
## Coeffecient and relevant statistics in dataframe

Get coeffecients etc.

If your right hand is on top when you clap, the odds are 14:1 that right is your writing hand rather than the left.

```{r}
model1 %>% 
  tidy(exponentiate = TRUE) %>%  # Transforms estimates into odds
  head()
```

## Variables for diagnostic check

Add fitted values and residuals to each observation

```{r}
model1_augmented <-
  model1 %>% 
  augment(type.predict = "response") %>% # Get fitted probabilities
  select(.fitted:.cooksd, everything()) # Reorder columns
head(model1_augmented)
```

TODO: Add some diagnostic plots / analysis of .cooksd (Dobson and Barnett, )

## Calibration plot

How well do fitted values correspond to observed proportions?

```{r}
model1_augmented %>% 
  mutate(W.Hnd_int = W.Hnd %>% as.integer() - 1) %>% 
  ggplot(aes(x = .fitted, y = W.Hnd_int, col = Sex)) + 
  geom_point(alpha = 0.3) + # Transparency of points
  geom_abline(slope =  1, 
              intercept = 0,
              linetype = "dashed") + 
  geom_smooth(se = FALSE) +  # loess smoother default
  coord_cartesian(xlim = c(0,1),
                  ylim = c(0,1)) +
  labs(x = "Predicted",
       y = "Observed")
```

Stratify further by exercise

```{r, fig.width=4, fig.height=6, fig.align='center'}
model1_augmented %>% 
  mutate(W.Hnd_int = W.Hnd %>% as.integer() - 1) %>% 
  ggplot(aes(x = .fitted, y = W.Hnd_int, col = Sex)) +
  facet_wrap(~Exer, ncol = 1) + 
  geom_point(alpha = 0.3) + # Transparency of points
  geom_abline(slope =  1, 
              intercept = 0,
              linetype = "dashed") + 
  geom_smooth(se = FALSE) +  # loess smoother default
  coord_cartesian(xlim = c(0,1),
                  ylim = c(0,1)) +
  labs(x = "Predicted",
       y = "Observed")
```



## Statistical tables for publication


```{r, results='asis'} 
# For html rendering
 # stargazer(model1, model2, 
 #           single.row = TRUE,
 #           type = "html",
 #           apply.coef = exp,
 #           header = FALSE,
 #            out = "test.html",
 #          report = "vc*")
```

```{r, results='asis'}
stargazer(model1, model2, 
          single.row = TRUE,
          type = "latex",
          apply.coef = exp,
          header = FALSE,
          report = "vc*")
```

Note that output can be saved in .tex and copied to latex

# For loops?

R and dplyr does not encourage the use of for loops (although it is possible).

```{r}
n = 0
for (i in c(1,2,3)) {
  n = i + 1
  print(n)
}
```


# Errors/debugging

Copy error message and google it.

# Input/output
